Reasoning about the rigor of perl programs: \\ \large gaps, limits to understanding, and means to ameliorate

### 2024-06-23 sciperl/paper.tex
more stuff

        note   = "TODO: chapter/verse"

@@

@@

The difficulty here is the scope of such an undertaking.
A lookup table of every subroutine in Perl's CORE\footnote{\texttt{perldoc perlfunc}} namespace's arguments mapping to perl's internal datatypes \footnote{see \texttt{perldoc perlguts}} would need to be built.
The same would have to be done for every FFI\footnote{Foreign Function Interface} or XS\footnote{\texttt{perldoc perlxs}} module used.
Given all other Perl subroutines you write are going to be a composition thereof, this knowledge would allow for correct characterizations of subroutines to be made.

@@

@@

Yet another place where I would love to have a proof of concept.

@@

TODO: show off POC doing this with tests...

>>>>

        note   = "ibid"

@@

    @misc{mltype,
        author = "Harry Mairson",
        year   = 1989,
        journal = "Proceedings of the ACM",
        url     = {https://dl.acm.org/doi/10.1145/96709.96748}
    @misc{clippert,
        title  = "Lambda Expressions vs. Anonymous Methods, pt. 5",
        url    = {https://learn.microsoft.com/en-us/archive/blogs/ericlippert/lambda-expressions-vs-anonymous-methods-part-five},
        author = "Eric Clippert",
        year   = 2007

@@

The difficulty here is twofold: the scope of such an undertaking and the computational difficulty to do so.
A lookup table of every subroutine in Perl's CORE\footnote{\texttt{perldoc perlfunc}} namespace's arguments mapping to perl's internal datatypes \footnote{see \texttt{perldoc perlguts}} would need to be built.
The same would have to be done for every FFI\footnote{Foreign Function Interface} or XS\footnote{\texttt{perldoc perlxs}} module used.
Given all other Perl subroutines you write are going to be a composition thereof, this knowledge would allow for correct characterizations of subroutines to be made.

@@

Nevertheless, this is a problem which others have attempted to solve and done so with some success.
Haskell, ML and Swift all have such type inference systems with the caveat of exponential worst-case execution times\cite{clippert}.
This is because the problem is in fact the 3SAT problem\footnote{also known as the boolean satisfiability problem} which is NP-hard\cite{mltype}.

Like all other NP-hard problems, they can be overcome by solving a simpler subset of the problem (aka "cheating").
A type inferencer can be lifted from one of these languages, and if we can devise a means to studiously avoid the perverse cases it would be of great utility.

@@

@@
